### 2017-02-07 doc/Paper.tex
Add model description

\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}

\usepackage{amsmath}  % equations
\usepackage{cite}  % bibliograby
\usepackage{graphicx}  % images
\usepackage{pgfplots}  % plots
  \pgfplotsset{width=10cm,compat=1.9} % plots settings
  \usepgfplotslibrary{external}
  \tikzexternalize

\usepackage{hyperref}  % references
  \hypersetup{colorlinks=true,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black} % references settings

\usepackage[automark]{scrpage2} % heading
  \pagestyle{scrheadings}
  \ihead[]{Handwritten character recognition in MPI} % left header
  \ohead[]{\today} % right header
  \cfoot[]{\pagemark} % footer
  \setheadsepline[122mm]{0.3mm}

\begin{document}
	\title{
	\Huge Handwritten digit recognition
	}

	\vspace{2cm}

	\author{\Large \href{mailto:stefan.niculae@my.fmi.unibuc.ro}{Stefan Niculae} \and \Large \href{mailto:ionut.ciocoiu@my.fmi.unibuc.ro}{Ionut Ciocoiu}
	\vspace{3cm}}

	\date{
	\large Parallel \& Concurrent Programming Lab Report \\
    \vspace{0.2cm}
	\today
	}

	\maketitle
	\setlength{\parindent}{0pt}

\vspace{5cm}
\begin{abstract}
We implement a neural network to classify handwritten digits. The training is done in parallel, using MPI. We provide a comparison for model hyper-parameters and for training time vs number of processes. It achieves great accuracy and fast training time. An interactive demo is also presented
\end{abstract}

\newpage
\tableofcontents

\newpage
\section{Problem statement}
\label{sec:problem}
Given a picture of a handwritten digit, label it accordingly 0 - 9.

\section{Dataset}
\label{sec:dataset}
We demonstrate our model performance on the MNIST database, widely used dataset for training and testing in the field of machine learning. It contains 70 000 examples of labeled handwritten digits. The digits have been size-normalized and centered in a fixed-size image \cite{mnist}.

\begin{figure}[htb]
\centering
\includegraphics[width=8cm]{images/mnist-images.png}
\caption{Sample MNIST images \cite{tf}}
\end{figure}

The images are black-and-white, with a resolution of $28 \times 28$ pixels, represented as an array of $784$ values. Each value ranges from $0$ to $1$ indicating the amount of blackness in the pixel.

\begin{figure}[htb]
\centering
\includegraphics[width=8cm]{images/mnist-matrix.png}
\caption{Matrix representation of a digit \cite{tf}}
\end{figure}

\newpage
\section{Model}
\label{sec:model}
We use a back-propagation neural network:
\begin{align*}
W^{(k)} &= W^{(k-1)} + \alpha \nabla_W^{(k-1)} \\ 
b^{(k)} &= b^{(k-1)} + \alpha \nabla_b^{(k-1)}
\end{align*}
Where $W$ is the weights matrix, $10 \times 784$ -- the number of classes( one for each digit) and the dimension of an image ($28 \times 28$). $b$ is the bias vector, $1 \times 10$, corresponding to the free element. $W^{(k)}$ denotes the weights at step $k$. $\nabla$ represents the update which must produce to improve prediction.
\\

The activation function is softmax:
$$f(x_i)=\frac{exp(x_i)}{\sum_j exp(x_j)}$$
Where $exp(x)$ represents $e^x$, Euler's constant. It \textit{normalizes} the vector, causing the greatest elements to have logarithmic relative value.
\\

The cost function is cross-entropy:
$$H(y_t, y_p)=-\sum_x y_t log(y_p)$$
Where $y_t$ is the true, correct label of an image and $y_p$ is the label predicted by the model.
\\

$L_2$ regularization is used:
$$\frac{1}{\lambda} \sum_i \sum_j W_{i,j}^2$$
Where $\lambda$ is the regularization parameter. This limits overfitting by imposing a penalty on big weights.
\\

The model uses an adaptive learning rate:
$$\eta \frac{ W^{(k)} - W^{(k-1)} } { \nabla^{(k)} - \nabla^{(k-1)} }$$
Where $\eta$ is the momentum parameter. This helps with stability in late epochs.

\begin{figure}[htb]
\centering
\includegraphics[width=8cm]{images/model-graph.png}
\caption{Simplified representation of model graph \cite{tf}}
\end{figure}

\paragraph{Training} We employ gradient-descent on mini-batches. Early-stopping prevents the model from further learning when performance on the validation split starts decreasing.

\paragraph{Evaluation}
TODO

\begin{tikzpicture}
\centering
\begin{axis} [
  title={Training performance},
  xlabel={Epoch},
  ylabel={Accuracy (\%)},
  xmin=0, xmax=100,
  ymin=0, ymax=100,
  xtick={0,25,50,75,100},
  ytick={0,25,50,75,100},
  legend pos=south east,
]

\addplot[]
  coordinates {
  (0,25)(25,65)(50,90)(100,95)
  };
\addplot[dashed]
  coordinates {
  (0,25)(25,50)(50,80)(100,90)
  };

\legend{train set, test set}
\end{axis}
\end{tikzpicture}

\newpage
\bibliographystyle{ieeetr}
	\bibliography{references} % expects file "references.bib"
	\addcontentsline{toc}{section}{References}

\end{document}